Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-10-07 03_marlin_paper.qmd
almost there

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.2"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

@@

## Need to cite..

@bosley2022

@berger2017

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
selectivity for individual species (each of which we denote as a *metier*). For
example, both a longline and purse seine fleet may capture bigeye (*Thunnus
obesus*) and skipjack tunas, but the longline fleet can be made much more likely
to catch larger bigeye than skipjack, and *vice versa*. Each fishing fleet
distributes its fishing effort in space according to a specified spatial
allocation rule (see *Spatial Allocation of Effort* ), for example in proportion
to profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

We use `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e. the
selected MPA design can be as patchwork as dictated by the data). We ran the
simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
set the area of each spatch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400
patches for a total area of 2,000 KM^2^.

@@

#### Pelagic Systems

@@

blue_climate_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  habitat = future_habitat,
  years = 20,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  ))
)

@@

For the coastal coral reef example both the *Spawning Ground* and *Target
fishing* MPA design strategies were capable of producing a Pareto-optimal or
improving, i.e. potentially win-win, outcome where change in total spawning
biomass across all species and yield across all fleets was greater than or equal
to zero for some MPA size. However, producing a total increase in SSB/SSB0
greater than 50% came at a net cost to fisheries yield across both design
strategies. The *Spawning Ground* strategy produced more immediate trade-offs
between food security and conservation (@fig-coral-results).

These macro-level changes in total yield and conservation are the result of
interesting internal dynamics that `marlin` allows users to explore through
output tables at the level of individual species and fleets. The *Spawning
Ground* strategy increased yield to Fleet One when MPAs were relatively small,
but was able to maintain yields near *status quo* levels up to a very large MPA
size to the dependence of Fleet One on only a relatively small portion of the
total simulation area. However, *Spawning Ground* MPAs were unable to provide
increases in yield to Fleet Two. The *Spawning Ground* strategy produced the
most rapid conservation gains for the snapper population. Interestingly, due to
effort displacement from the spawning grounds onto the fishing areas inhabited
by the groupers the rest of the year, closure of the spawning ground had limited
impact on the grouper population until the MPA size exceeded 50% of the
simulation area. Effort displacement from the spawning grounds also resulted in
a small decrease in the deep snapper population for MPA sizes less than 50%. The
*Target Fishing* design strategy has similar yield outcomes to the *Spawning
Ground* strategy, but less dramatic negative impacts on Fleet Two, at the
expense of slightly fewer benefits for Fleet One. However, the *Target Fishing*
strategy provided more uniform conservation benefits to all species
(@fig-coral-results).

MPAs were capable of producing positive outcomes in terms of total biomass and
yields in the coastal coral reef example, but all the tested strategies resulted
in trade-offs across at least one dimension. `marlin` could be further used by a
community such as the one simulated here to explore alternative design
strategies, or to assess options for distributing fishery benefits across
communities that may experience very different food provision outcomes from the
same MPA network.

@@

@@

coral_frontier <- tmp %>%
  filter(prop_mpa < 0.7) |>
  ggplot(aes(delta_biodiv, delta_yield)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_path(alpha = 0.8, aes(color = placement_strategy),size = 1.1) +
  geom_point(aes(fill = prop_mpa), shape = 21, size = 3) +
  # geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2), limits = c(0,0.3)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  # guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) +
  scale_fill_viridis(name = "% MPA", labels = scales::label_percent(accuracy = 1),guide = guide_colorbar(ticks.colour = "black",frame.colour = "black"), option = "cividis") +
  labs(tag = "C")  +
  theme(
    legend.position = c(0.73, .73),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 8)
  )
coral_fleet_yield <- coral_results %>%
  pivot_longer(starts_with("fleet_"),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
  ungroup() %>% 
  mutate(fleet= fct_relabel(fleet, titler)) %>% 
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Yield",
                     labels = scales::label_percent(accuracy = 1)) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        legend.margin = margin())  + 
    scale_color_discrete(name = '') + 
    labs(tag = "A")

@@

### Pelagic Systems

The yield impacts of the proposed MPA strategies varied across each of our
simulations. Yields of the purse seine fleet were more resilient to the range
shift under the *Rate* design strategy, while the longline fleet had the exact
opposite result. Under the *Rate* strategy, longline yields were relatively
stable across a large MPA size under the *status quo* conditions, but declined
rapidly under the *range shift* conditions. For the purse seine fleet, yields
were more stable under the *range shift* conditions under the *Target Fishing*
design scenario, but declined quickly as a function of MPA size under the
*status quo* conditions (@fig-blue-results). In terms of fisheries yield, the
*Target Fishing* strategy would produce better yield outcomes for both fishing
fleets under the *range shift* conditions, but the *Rate* strategy would perform
best under the *status quo*.

From a conservation perspective, the primary tuna species (bigeye, skipjack, and
yellowfin) were most sensitive to design strategy and climate scenario, with the
*Target Fishing* strategies producing for example rapid conservation gains for
bigeye under the *status quo*, while the *Rate* strategy resulted in net
conservation losses until the MPA became extremely large. However, under the
*Range Shift* scenario the MPAs had little impact on many of the tuna
populations until the MPA size became extremely large, do to movement of the
primary fishing grounds outside of the current hotpots. The 20 year time horizon
simulated here was not enough to produce substantial gains for any of the shark
species even with 100% closures except for the more rapidly growing blue shark
(@fig-blue-results).

@@

xx add more details on case study here xx

-   Coral

    -   targeting spawning grounds for MPA provides better / more equitable
        yield outcomes, but at the expense of much slower conservation gains,
        and even some negative tradeoffs for species like deep snapper.
        Targeting fishing provides better gains to fleet two (explain why), and
        better conservation outcomes, but squarely at the expense of fleet one

-   Blue Water

    -   target-fishing design being much more robust to range shifts than rate
        design, why?

    -   High concentration of the purseseine fleet means much more sensitive to
        placement "errors"

    -   Basic life history poses challenge to recovery of severely depleted
        species e.g. OWT

@@

@@

Marine social-ecological systems are highly complex. The model presented here is
designed to help users explore the impacts of different variables on policy
outcomes. But, that freedom comes at a cost, namely that users have a large
number of options at their disposal that they must decide on. Our recommended
strategy for dealing with this is for users to try and narrow down a list of
parameters that they feel are sufficiently "known" and another list of
parameters that are highly uncertain and / or contentious which the users feel
may impact results.

For example, a community seeking to a model a well-studied coastal coral reef
ecosystem might leave as fixed the habitat distribution (as represented by reef
locations) and general life history of the species in question (growth rates,
age at maturity, etc). But, within those constraints, the users may wish to test
the sensitivity of proposed MPA networks to key unknowns such as adult and
larval dispersal rates or the the economic incentives and gear selectivity of
the fishing fleets. As an example of this, we ran an alternative version of our
Coral Reef case study in which Fleet Two was assigned logistic selectivity for
the snapper and grouper species, rather than the dome-shaped selectivity
presented in our main results. We found that the impacts of MPAs on conservation
and food security were relatively insensitive to the form of fishing selectivity
(Fig.S1). This does not mean that knowing the "correct" form of selectivity may
not be extremely important in providing accurate assessment results
[@waterhouse2014], but rather that holding other parameters constant, in this
case misspecifying the selectivity curve is projected to have little impact on
the simulated outcomes of MPAs.

@@

`marlin` is just one model out of many that can be used to inform marine
management. Other models may be better suited to address specific forms of
complexity, such as predatory-prey dynamics (citations) or more complex fleet
dynamics through agent-based approached [@burgess2020]. Particularly for more
tactical applications, we would encourage users to explore multiple modeling
frameworks to help design policies that are likely to be robust to many
different kinds of complexities.

In sum, `marlin` complements the existing spatial marine modeling literature by
allowing scientists, decision makers, and other stakeholders to efficiently
examine the impacts of realistic bio-economic dynamics on academic and applied
problems. For example, scientists could use `marlin` to consider the relative
importance of different biological variables in affecting policy outcomes.
Managers could use `marlin` to assess the potential of proposed MPA networks to
achieve stated objectives, such as maximizing conservation gains for a
threatened species while minimizing or removing negative impacts to affected
fishing fleets.

Beyond the current examples, we envision `marlin` being applicable to research
on dynamic ocean management, range shifts, management strategy evaluation,
policy interactions, and spatial stock assessment. In addition, the
process-based movement model used in `marlin` can directly use empirical
estimates of movement dynamics derived in the manner of @thorson2021a, providing
a link between empirical and simulation based approaches to marine resource
management. The model currently focuses on diffusion and taxis, which allows for
general representation of animals following physical or oceanographic features,
but research on the incorporation and importance of advective forces would be of
value going forward. Future iterations of the model may include life history
modules for a broader range of species types such as marine mammals, sea birds,
and sea turtles.

@@

This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for
helpful revisions of this manuscript.

>>>>

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.1"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

@@

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
Each fishing fleet distributes its fishing effort in space according to a
specified spatial allocation rule (see *Spatial Allocation of Effort* ), for
example in proportion to profit-per-unit-effort, conditional on management
policies in places such as quotas and/or the presence of any spatial
restrictions such as MPAs. It is important to note that contact selectivity
reflects the ability of the fishing method in question to capture fish of
different sizes that come into "contact" with the gear. The contact selectivity
of each of the specified *metiers* will then interact with the distribution of
fish sizes and fishing effort in space simulated by `marlin` to produce a net
"population" selectivity, which may differ from the individual contact
selectivities of each of the *metiers*.

@@

We use `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e. the
selected MPA design can be as patchwork as dictated by the data). We ran the
simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
patches for a total area of 2,000 KM^2^.

@@

#### Pelagic Ecosystems

@@

blue_climate_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  habitat = future_habitat,
  years = length(future_habitat[[1]]),
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  ))
)

@@

```{r}
#| label: coral-results

@@

coral_fleet_results <- coral_results %>%
  pivot_longer(starts_with("fleet_"),
               names_to = "fleet",
               values_to = "fleet_yield") %>%
  group_by(prop_mpa, fleet, placement_strategy) %>%
  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
  group_by(fleet, placement_strategy) %>%
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) |> 
  ungroup()

peak_coral_fleet_results <- coral_fleet_results |> 
  group_by(fleet) |> 
  filter(delta_yield == max(delta_yield)) |> 
  mutate(peak_yields = scales::percent(delta_yield))

peak_coral_fleet_results |> filter(fleet == "fleet_two")

MPAs were capable of producing a range of positive and negative outcomes for
food security and conservation in the Coral Reef case study depending on the
design strategy used and the size of MPA implemented. Both of the MPA design
strategies were capable of increasing fisheries yield for Fleet Two, up to a
value of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
even when MPAs covered more than 50% of the simulated area. However, Fleet One
only benefited from MPAs under the *Spawning Ground* design strategy, with a
maximum increase of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
the *Target Fishing* design strategy produced a roughly linear decrease in
fishing yields as a function of increasing MPA size (@fig-coral-results A).

On the conservation front, MPAs were uniformly beneficial to the spawning
biomass of all species under the *Target Fishing* design strategy, with the most
rapid increases for the Deep Snapper population. The *Spawning Ground* strategy
primarily benefited the coastal Snapper population, producing little change in
the Grouper population and decreasing spawning biomass of both the Deep Snapper
and Reef Shark populations for MPA sizes covering less than 50% of the simulated
area (@fig-coral-results B).

In total, the *Spawning Ground* strategy was capable or providing net increases
in fishing yield (summed across both fleets) of up to roughly 7%, with positive
net yield impacts up to network size of nearly 60% of the area. However, even
MPAs covering 60% of the area only produced a maximum increase of roughly 5% in
total SSB/SSB0 when designed around the spawning ground. Conversely, the *Target
Fishing* design strategy was capable of producing a nearly 25% increase in total
SSB/SSB0 for the same MPA size, but at a much greater cost to the total food
production from the system's fisheries (@fig-coral-results C).

<!-- ~~These macro-level changes in total yield and conservation are the result of -->

<!-- interesting internal dynamics that marlin allows users to explore through output -->

<!-- tables at the level of individual species and fleets. The *Spawning Ground* -->

<!-- strategy increased yield to Fleet One when MPAs were relatively small, but was -->

<!-- able to maintain yields near *status quo* levels up to a very large MPA size to -->

<!-- the dependence of Fleet One on only a relatively small portion of the total -->

<!-- simulation area. However, *Spawning Ground* MPAs were unable to provide -->

<!-- increases in yield to Fleet Two. The *Spawning Ground* strategy produced the -->

<!-- most rapid conservation gains for the snapper population. Interestingly, due to -->

<!-- effort displacement from the spawning grounds onto the fishing areas inhabited -->

<!-- by the groupers the rest of the year, closure of the spawning ground had limited -->

<!-- impact on the grouper population until the MPA size exceeded 50% of the -->

<!-- simulation area. Effort displacement from the spawning grounds also resulted in -->

<!-- a small decrease in the deep snapper population for MPA sizes less than 50%. The -->

<!-- *Target Fishing* design strategy has similar yield outcomes to the *Spawning -->

<!-- Ground* strategy, but less dramatic negative impacts on Fleet Two, at the -->

<!-- expense of slightly fewer benefits for Fleet One. However, the *Target Fishing* -->

<!-- strategy provided more uniform conservation benefits to all species -->

<!-- (@fig-coral-results).~~ -->

@@

coral_frontier <- tmp %>%
  filter(prop_mpa < 0.7) |>
  ggplot(aes(delta_biodiv, delta_yield)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_path(alpha = 0.8, aes(color = placement_strategy),size = 1.1) +
  geom_point(aes(fill = prop_mpa), shape = 21, size = 3) +
  # geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2), limits = c(0,0.3)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  # guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) +
  scale_fill_viridis(name = "% MPA", labels = scales::label_percent(accuracy = 1),guide = guide_colorbar(ticks.colour = "black",frame.colour = "black"), option = "cividis") +
  labs(tag = "C")  +
  theme(
    legend.position = c(0.73, .73),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 8)
  )
coral_fleet_yield <- coral_fleet_results |> 
  mutate(fleet= fct_relabel(fleet, titler)) %>% 
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Yield",
                     labels = scales::label_percent(accuracy = 1)) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        legend.margin = margin())  + 
    scale_color_discrete(name = '') + 
    labs(tag = "A")

@@

### Pelagic Ecosystems

The yield impacts of the proposed MPA strategies varied across each of our
simulations. Yields of the purse seine fleet were more resilient to the range
shift under the *Rate* design strategy, while the longline fleet had the exact
opposite result. Under the *Rate* strategy, longline yields were relatively
stable across a large MPA size under the *status quo* conditions, but declined
rapidly under the *range shift* conditions. For the purse seine fleet, yields
were more stable under the *range shift* conditions under the *Target Fishing*
design scenario, but declined quickly as a function of MPA size under the
*status quo* conditions (@fig-blue-results A). In terms of fisheries yield, the
*Target Fishing* strategy would produce better yield outcomes for both fishing
fleets under the *range shift* conditions, but the *Rate* strategy would perform
best under the *status quo*.

From a conservation perspective, the primary tuna species (bigeye, skipjack, and
yellowfin) were most sensitive to design strategy and climate scenario, with the
*Target Fishing* strategies producing for example rapid conservation gains for
bigeye under the *status quo*, while the *Rate* strategy resulted in net
conservation losses until the MPA became extremely large. However, under the
*Range Shift* scenario the MPAs had little impact on many of the tuna
populations until the MPA size became extremely large, do to movement of the
primary fishing grounds outside of the current hotpots. The 20 year time horizon
simulated here was not enough to produce substantial gains for any of the shark
species even with 100% closures except for the more rapidly growing blue shark
(@fig-blue-results B).

@@

Examining our coral reef case study, our results indicate how this model could
be used to examine the sensitivity of conservation and food security outcomes of
MPAs to alternative design scenarios, holding constant other social-ecological
variables. The *Spawning Ground* MPA network was able to provide more equitable
yield outcomes across the two fishing communities, whereas the *Target Fishing*
MPA network only benefited Fleet Two (@fig-coral-results A). This is because
Fleet One has two primary fishing grounds; the spawning grounds, and the
offshore area where the deep snapper live. Fleet Two primarily fishes in the
north west corner of the simulation grid. Under the *Spawning Ground* strategy,
while Fleet One quickly losses fishing grounds on the spawning grounds, it can
compensate for this by moving offshore and fishing the deep snapper population
harder, while also being able to fish some of the spillover of snapper from the
spawning ground closure. This is why biomass of deep snapper actually declines
under the *Spawning Ground* strategy.

Conversely, most of the catch in the fishery comes from the nearshore and deep
water snapper populations, which overlaps with both of Fleet One's fishing
grounds. The *Target Fishing* strategy then begins closing both the nearshore
and offshore snapper fishing grounds, resulting in too much of a loss in fishing
grounds for Fleet One to be offset by spillover from the MPAs. As a result of
primarily placing MPAs on Fleet One's fishing grounds, Fleet Two gains spillover
benefits at little cost to their fishing grounds until the MPAs reach their
fishing grounds once protection nears 100%. By protecting both the spawning
ground and the offshore areas, the *Target Fishing* network provides
conservation benefits to all of the species, in contrast to the *Spawning
Ground* strategy that only provides meaningful conservation gains to the snapper
population up until very large MPA sizes. Our result that protection of a
dedicated spawning ground did not provide substantially greater conservation
outcomes for the species using that spawning ground (groupers and sharks) than
an alternative design strategy is supported by other modeling studies that show
that the impacts of spawning ground protection on conservation and yields may be
highly variable, and that displacement of high levels fishing effort from the
spawning grounds can offset potential conservation gains of the protection
[@grüss2014] .

The two design strategies produced very different outcomes in terms of total
changes in yield and conservation outcomes (@fig-coral-results C). Both MPA
network design strategies were capable of producing net "win-win" outcomes in
which both food security (as measured by change in total fisheries yields) and
conservation (as measured by change in SSB/SSB0). However, in general for the
same size MPA network the *Spawning Ground* design strategy produced better food
security outcomes but worse conservation outcomes, and *vice* *versa* for the
*Target Fishing* strategy. The coral reef case study shows how the modeling
framework presented here can help stakeholders explore how different management
strategies affect outcomes both in total (@fig-coral-results C) and across
species (@fig-coral-results A) and fishing communities (@fig-coral-results B).

Turning to our Pelagic Ecosystem case study, we see that overall the yield
outcomes of the purse-seine fleet were much more sensitive to the presence of a
range shift than the longline fleet, particularly under the *Target Fishing*
design strategy. This is because the purse-seine fleet primarily targets
skipjack tuna, which in this simulation are concentrated in a relatively narrow
latitudinal band (@fig-blue), and the purse-seine fleet makes up a large portion
of the total catch in the simulated fishery. So, the "Target Fishing" strategy
starts by closing off the main purse-seine fishing grounds, which while only
small part of the spatial domain of the model represents a large portion of the
purse-seine fleet's fishing grounds, resulting both in more rapid conservation
gains and fishery losses under the *Status Quo* species distributions. The
*Rate* strategy places more MPAs in areas that are of lesser importance fo the
purse-seine fleet but overlap more with species such as Oceanic Whitetip Shark.
Conversely, the purse-seine fleet appears to do have better MPA yield outcomes
under the Range Shift scenario, not because of rebuilding of the Skipjack
population, but due to future fishing grounds being essentially unprotected by
smaller MPAs targeting the current skipjack distribution ( @fig-blue-results ).

The conservation outcomes of the skipjack, bigeye, and yellowfin tunas were
among the most sensitive to range shifts, whereas even complete closure of the
region was not sufficient to see significant increases in severely depleted
species like oceanic whitetip sharks within the 20-year timeline of the
simulation. Future research could be conducted then to see what sorts of
timelines might be needed to see recovery of these species, and which design
strategies result in the fastest recovery at the lowest cost to other objectives
such as food security.

These case studies are only intended to illustrate the capabilities of the model
and the importance of considering the bio-economic dynamics represented in
`marlin`; it is beyond the scope of this paper to provide broader conclusions
about the performance of MPAs under different contexts or a comprehensive
comparison of simulation and empirical results of MPAs around the world. That
being said, the kinds of dynamics resulting in our case studies are well
supported by both modeling and empirical studies. Our results support conceptual
[@hilborn2004a; @gaines2010] and empirical [@ban2019] evidence that under the
right conditions no-take MPAs can benefit fisheries and conservation.
@rassweiler2014 demonstrated that the kinds of design choices presented in our
case studies can greatly drive MPA outcomes. @ovando2021a provides empirical
support for the possibility of even well enforced and large MPAs to have less
pronounced population-level conservation effects (@fig-coral-results A) than
might be expected given observed differences in biomass inside and outside of
protected areas.

MPAs based around coral-reef style ecosystems have been extensively studied
around the world, with much of the empirical evidence of their performance
centered on comparing measures such as biomass densities inside protected areas
relative to outside [@lester2009]. Our model predicts a similar rapid increase
of the simple ratio of mean biomass inside MPAs relative to outside (Fig.XX,
acknowledging that designing a proper response ratio would require controlling
for habitat characteristics and MPA design criteria). However, despite producing
clear response ratios, The net conservation and fishery outcomes of our
simulated coral-reef ecosystem MPAs were not nearly as large or clear as the
response ratio results (@fig-coral-results); this result is supported works such
as @ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
indicator of the net causal impacts of MPAs.

Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
empirical support for our case study results showing how MPAs can produce
trade-offs across the conservation of different species and fleets when the
spatial-temporal distributions and life history traits of affected species are
heterogeneous, and MPAs result in some degree of effort displacement or
concentration in the remaining fishing grounds. @hampton2023 supports our
pelagic case study result showing that both conservation and fishery impacts of
MPAs on the highly mobile species of the open oceans can be limited unless
protected areas are very large. @brown2018 supports our result showing how
heterogeneity in the behavior and objectives of fishing fleets sharing an
ecosystem can affect the magnitude and equity of fishery reform outcomes.
@davies2017 discusses the importance of considering the potential of climate
driven range shifts when assessing spatial management policies; our work builds
on this by allowing users to not just simulate species distributions but also
distribution of biomass and age composition of fish space and time under climate
change.

@@

`marlin` does not represent trophic interactions. Anthropgenic changes in
species abundance can clearly result in trophic cascades throughout the affected
food web. However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
first decades of protection. Signals of trophic cascades may be masked by
variations in the direction and strength of species interactions driven by
environmental context [@liu2022], or may simply take longer to develop
detectable effects than the coverage of many time series of MPAs. This is to say
then that while `marlin` does not incorporate trophic interactions, what
empirical evidence we have does not suggest that management-mediated trophic
cascades are so common and clear that they must be incorporated into any
credible multi-species simulation model. Rather, there is still value in
capturing other dynamic bio-economic processes, as done here, even if that comes
at the expense of modeling trophic interactions. However, research on the
trophic impacts of spatial-temporal management actions is clearly of value and
models like Atlantis [@audzijonyte2019], Ecopath with Ecosim [@christensen2004],
and EASI-Fish [@griffiths2019] can help users explore those kinds of processes,
in the manner of @baskett2007.

@@

Marine social-ecological systems are highly complex. The model presented here is
designed to help users explore the impacts of different variables on policy
outcomes. But, that freedom comes at a cost, namely that users have a large
number of options at their disposal that they must decide on. Our recommended
strategy for dealing with this is for users to try and narrow down a list of
parameters that they feel are sufficiently "known" and another list of
parameters that are highly uncertain and / or contentious which the users feel
may impact results. Where possible, parameters from locally estimated stock
assessment models can be used to provide a foundation around which sensitivity
analyses around specific uncertainties of interest [@berger2017].

For example, a community seeking to a model a well-studied coastal coral reef
ecosystem might leave as fixed the habitat distribution (as represented by reef
locations) and general life history of the species in question (growth rates,
age at maturity, etc). But, within those constraints, the users may wish to test
the sensitivity of proposed MPA networks to key unknowns such as adult and
larval dispersal rates or the the economic incentives and gear selectivity of
the fishing fleets. As an example of this, we ran an alternative version of our
Coral Reef case study in which Fleet Two was assigned logistic selectivity for
the snapper and grouper species, rather than the dome-shaped selectivity
presented in our main results, reflecting a scenario where users are uncertain
about the contact selectivity of Fleet Two. We found that the impacts of MPAs on
conservation and food security were relatively insensitive to the form of
contact selectivity specified for Fleet Two (Fig.S1). This does not mean that
knowing the "correct" form of contact selectivity may not be extremely important
in providing accurate assessment results [@waterhouse2014], but rather that in
this case misspecifying the contact selectivity curve is projected to have
little impact on the simulated outcomes of MPAs, conditional on holding other
parameters constant.

@@

`marlin` is just one model out of many that can be used to inform marine
management. Other models may be better suited to address specific forms of
complexity, such as predatory-prey dynamics (citations) or more complex fleet
dynamics through agent-based approached [@burgess2020a]. Particularly for more
tactical applications, we would encourage users to explore multiple modeling
frameworks to help design policies that are likely to be robust to many
different kinds of complexities.

At this time, `marlin` assumes that all evaluated policies are perfectly
implemented; e.g. that MPAs are 100% no-take and perfectly enforced, that quotas
and closed seasons are respected, that there is no discard mortality, etc. In
reality no policy is perfectly implemented, and users should consider the extent
to which the policies they simulate are actually feasible to implement. There is
value though in being able to simulate and compare the outcomes of perfectly
implmented policies to isolate the concept of the polucy itself from its
implementation. That being said, future work may add the capability for
imperfect implmentation to the model.

`marlin` complements the existing spatial marine modeling literature by allowing
scientists, decision makers, and other stakeholders to efficiently examine the
impacts of realistic bio-economic dynamics on academic and applied problems. For
example, scientists could use `marlin` to consider the relative importance of
different biological variables in affecting policy outcomes. Managers could use
`marlin` to assess the potential of proposed MPA networks to achieve stated
objectives, such as maximizing conservation gains for a threatened species while
minimizing or removing negative impacts to affected fishing fleets.

Beyond the current examples, we envision `marlin` being applicable to research
on dynamic ocean management, range shifts, management strategy evaluation,
policy interactions, and spatial stock assessment. `marlin` can help researchers
generate data for further testing of the performance and design of spatially
explicit integrated population models, in the manner of @bosley2022. In
addition, the process-based movement model used in `marlin` can directly use
empirical estimates of movement dynamics derived in the manner of @thorson2021a,
providing a link between empirical and simulation based approaches to marine
resource management that has been challenging to implement in spatial
simulations [@berger2017]. The model currently focuses on diffusion and taxis,
which allows for general representation of animals following physical or
oceanographic features, but research on the incorporation and importance of
advective forces would be of value going forward. Future iterations of the model
may include life history modules for a broader range of species types such as
marine mammals, sea birds, and sea turtles.

Fisheries models, assessments, management have often abstracted away many of the
spatial-temporal complexities of marine social-ecological systems [@ovando2021;
@berger2017]. The modeling framework described here can help facilitate the
science and application of spatial fisheries management by supporting the
simulation of different spatial-temporal dynamics to aid in testing of various
aspects of the marine resource management process.

@@

This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston and
two anonymous reviewers for helpful revisions and comments of this manuscript.
